PhD Thesis

### 2017-01-20 pldi14/sec-concl.tex
Drop duplicated label
Revise ToC for second part
Move sections to chapters and so on.

\section{Conclusions and future work}
\label{sec:concl}
\label{ssec:future}
We have presented \ILC, an approach to lifting incremental computations
on first-order programs to incremental computations on higher-order
programs. We have presented a machine-checked correctness proof 
of a formalization of \ILC\ and an initial experimental evaluation
in the form of an implementation, a sample plugin for maps and bags,
and a non-trivial example that was incrementalized successfully and
efficiently.

Our work opens several avenues of future work. Our current implementation
is not efficient on derivatives that are not self-maintainable.
However, as discussed
(Sec.~\ref{ssec:self-maint}), we will study how
to memoize intermediate results to address this limitation. Our next
step will be to develop language plugins which
have efficient non-self-maintainable primitives.

>>>>

\section{Chapter conclusion}
\label{ssec:future}
In this chapter we have presented \ILC, an approach to lifting incremental computations
on first-order programs to incremental computations on higher-order
programs. We have presented a machine-checked correctness proof 
of a formalization of \ILC\ and an initial experimental evaluation
in the form of an implementation, a sample plugin for maps and bags,
and a non-trivial example that was incrementalized successfully and
efficiently.

\pg{Revise}
Our work opens several avenues of future work. Our current implementation
is not efficient on derivatives that are not self-maintainable.
However, as discussed
(Sec.~\ref{ssec:self-maint}), we will study how
to memoize intermediate results to address this limitation. Our next
step will be to develop language plugins which
have efficient non-self-maintainable primitives.

### 2017-01-20 pldi14/sec-differentiate.tex
Split fig:base-calculus into separate figures
This avoids the overlap between the two subfigures.
Revise ToC for second part
Move sections to chapters and so on.

\section{Incrementalizing \TitleLambda{}-calculi}
\label{sec:differentiate}
\label{sec:correctness}

@@

\pg{Here we need an example like the one in Sec. 2.2, using the syntactic counterpart to closures, that is open terms.}
In this section, we show how to incrementalize an arbitrary
program in simply-typed $\Gl$-calculus. To this end, we define
the source-to-source transformation $\DERIVE$. Using the
denotational semantics $\Eval{-}$ we define later (in
\cref{sec:denotational-sem}), we can specify $\DERIVE$'s intended
behavior: to ensure \cref{eq:correctness},
$\Eval{\Derive{f}}$ must be the derivative of $\Eval{f}$
for any closed term $\HasType{f}{A \to B}$. We will overload the word
``derivative'' and say simply that $\Derive{f}$ is the derivative of
$f$.

It is easy to define derivatives of arbitrary functions as:
\[\App{\App{f'}{x}}{\D x} = \Diff{\App{f}{\Update*{x}{\D x}}}{\App f x}\text{.}\]
We could implement $\DERIVE$ following the same strategy.
However, the resulting incremental programs would be no faster
than recomputation. We cannot do better for arbitrary mathematical functions,
since they are infinite objects which we cannot fully inspect.
%
\pg{Revise - I reordered sentences to sort them logically with minimal rewriting.}%
Therefore, we resort to a source-to-source transformation
on simply-typed $\Gl$-calculus as defined in 
\cref{fig:base-calculus}. In this section, we focus on the
incrementalization of the features that are shared among all
instances of the plugin interface, that is, function types and the
associated syntactic forms, $\Gl$-abstraction, application and
variable references.

@@

\subsection{Change types and erased change structures}
\label{ssec:change-types}

@@

\subsection{Differentiation}
\label{ssec:differentiation}
% The following should explain the invariant for \DERIVE.

@@

\subsection{Plugins}\label{ssec:plugin}

>>>>

\chapter{Static differentiation for simply-typed \TitleLambda{}-calculus}
%\section{Incrementalizing \TitleLambda{}-calculi}
\label{sec:differentiate}
\label{sec:correctness}

@@

\pg{Here we need an example like the one in Sec. 2.2, using the syntactic counterpart to closures, that is open terms.}
In this chapter, we show how to incrementalize an arbitrary
program in simply-typed $\Gl$-calculus. To this end, we define
the source-to-source transformation $\DERIVE$. Using the
denotational semantics $\Eval{-}$ we define later (in
\cref{sec:denotational-sem}), we can specify $\DERIVE$'s intended
behavior: to ensure \cref{eq:correctness},
$\Eval{\Derive{f}}$ must be the derivative of $\Eval{f}$
for any closed term $\HasType{f}{A \to B}$. We will overload the word
``derivative'' and say simply that $\Derive{f}$ is the derivative of
$f$.

It is easy to define derivatives of arbitrary functions as:
\[\App{\App{f'}{x}}{\D x} = \Diff{\App{f}{\Update*{x}{\D x}}}{\App f x}\text{.}\]
We could implement $\DERIVE$ following the same strategy.
However, the resulting incremental programs would be no faster
than recomputation. We cannot do better for arbitrary mathematical functions,
since they are infinite objects which we cannot fully inspect.
%
\pg{Revise - I reordered sentences to sort them logically with minimal rewriting.}%
Therefore, we resort to a source-to-source transformation
on simply-typed $\Gl$-calculus as defined in 
\cref{fig:syntax,fig:typing}. In this section, we focus on the
incrementalization of the features that are shared among all
instances of the plugin interface, that is, function types and the
associated syntactic forms, $\Gl$-abstraction, application and
variable references.

@@

\section{Change types and erased change structures}
\label{ssec:change-types}

@@

\section{Differentiation}
\label{ssec:differentiation}
% The following should explain the invariant for \DERIVE.

@@

\section{Plugins}\label{ssec:plugin}

### 2017-01-20 pldi14/sec-practice.tex
Revise ToC for second part
Move sections to chapters and so on.

% Emacs, this is -*- latex -*-!
\section{Differentiation in practice}
\label{sec:applying}

@@

\subsection{The role of differentiation plugins}
\label{ssec:methodology}
Users of our approach need to
%
(1) choose which base types and primitives they need,
%
(2) implement suitable differentiation plugins for these base
types and primitives,
%
(3) rewrite (relevant parts of) their programs in terms of these primitives and
%
(4) arrange for their program to be called on changes instead of
updated inputs.

@@

  \begin{oldSec}
\subsection{Methodology}
\label{ssec:methodology}

@@

\subsection{Predicting nil changes}
Handling changes to all inputs can induce excessive overhead in incremental
programs~\citep{Acar09}. It is also often
unnecessary; for instance, the function argument of $\Term{fold}$ in 
\cref{sec:intro} does not change since it is a closed subterm of the program, so
$\Term{fold}$ will receive a nil change for it.
A (conservative) static analysis can detect changes that are guaranteed to be nil at runtime. 
We can then specialize derivatives that receive this change, so that they need not
inspect the change at runtime.

@@

\subsection{Case study}
\label{sec:plugins}

@@

\begin{oldSec}
\subsubsection{Bags}

@@

\subsubsection{Maps}

@@

\subsection*{Details}

@@

Moreover, we can show that those axioms are consistent by
providing models implementing them. \pg{Resume}
%Additionally, we have two implementations of those axioms.
\pg{A couple more axioms need to be checked.}
%
% (s₂ ⊝ (ds ⊕ s₁)) ∘ ds &= s₂ ⊝ s₁ (?)
%
% Special case:
%	(ds ⊕ s) ⊝ s = ds (EDIT: only if ds is valid for s).
%
% Therorems:
%
%s ⊝ s = nil s
%(s₁ ⊝ s₂) ∘ (s₂ ⊝ s₃) = s₁ ⊝ s₃
\end{oldSec}
\subsection{Benchmarking setup}
We run object language programs by generating corresponding Scala code.
To ensure rigorous
benchmarking~\citep{Georges07rigorousJavaPerformance}, we use the
Scalameter benchmarking library. To show that the performance
difference from the baseline is statistically significant, we
show 99\%-confidence intervals in graphs.

@@

\subsection{Benchmark results}

>>>>

% Emacs, this is -*- latex -*-!
\chapter{Differentiation in practice}
\label{sec:applying}

@@

\section{The role of differentiation plugins}
\label{ssec:methodology}
Users of our approach need to
%
(1) choose which base types and primitives they need,
%
(2) implement suitable differentiation plugins for these base
types and primitives,
%
(3) rewrite (relevant parts of) their programs in terms of these primitives and
%
(4) arrange for their program to be called on changes instead of
updated inputs.

@@

  \begin{oldSec}
\section{Methodology}
\label{ssec:methodology}

@@

\section{Predicting nil changes}
Handling changes to all inputs can induce excessive overhead in incremental
programs~\citep{Acar09}. It is also often
unnecessary; for instance, the function argument of $\Term{fold}$ in 
\cref{sec:intro} does not change since it is a closed subterm of the program, so
$\Term{fold}$ will receive a nil change for it.
A (conservative) static analysis can detect changes that are guaranteed to be nil at runtime. 
We can then specialize derivatives that receive this change, so that they need not
inspect the change at runtime.

@@

\section{Case study}
\label{sec:plugins}

@@

\begin{oldSec}
\subsection{Bags}

@@

\subsection{Maps}

@@

\section*{Details}

@@

Moreover, we can show that those axioms are consistent by
providing models implementing them. \pg{Resume}
%Additionally, we have two implementations of those axioms.
\pg{A couple more axioms need to be checked.}
%
% (s₂ ⊝ (ds ⊕ s₁)) ∘ ds &= s₂ ⊝ s₁ (?)
%
% Special case:
%	(ds ⊕ s) ⊝ s = ds (EDIT: only if ds is valid for s).
%
% Therorems:
%
%s ⊝ s = nil s
%(s₁ ⊝ s₂) ∘ (s₂ ⊝ s₃) = s₁ ⊝ s₃
\end{oldSec}
\section{Benchmarking setup}
We run object language programs by generating corresponding Scala code.
To ensure rigorous
benchmarking~\citep{Georges07rigorousJavaPerformance}, we use the
Scalameter benchmarking library. To show that the performance
difference from the baseline is statistically significant, we
show 99\%-confidence intervals in graphs.

@@

\section{Benchmark results}
